---
title: "Gower Exponent Examples"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    theme: cerulean
    highlight: textmate
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE)
```

This document contains an example that demonstrates the effect an exponent has on the similarity computation between a prediction of interest and another observation.

```{r}
# Load libraries
library(dplyr)
library(ggplot2)
library(tidyr)
```

In the *lime* R package, the similarity between the prediction of interest (*poi*) and another observation (*obs*) is computed as:

> 1 - (Gower distance(*poi*, *obs*)^(exponent))

where the "exponent" is specified by the user.

The code below creates an example dataset that computes similarities based on various exponents from the same distances.

```{r}
# Prepare an example dataset
dist_data <-
  data.frame(distance = seq(0, 1, length.out = 100)) %>%
  mutate(
    sim0.1 = 1 - ((distance) ^ 0.1),
    sim0.25 = 1 - ((distance) ^ 0.25),
    sim0.5 = 1 - ((distance) ^ 0.5),
    sim1 = 1 - ((distance) ^ 1),
    sim10 = 1 - ((distance) ^ 10),
    sim15 = 1 - ((distance) ^ 15),
    sim5 = 1 - ((distance) ^ 5)
  ) %>%
  pivot_longer(names_to = "exponent",
               values_to = "similarity",
               cols = -distance) %>%
  mutate(exponent = stringr::str_remove(exponent, "sim")) %>%
  mutate(exponent = factor(
    exponent,
    levels = c("0.1", "0.25", "0.5", "1", "5", "10", "15")
  ))

# Print the head of the data
head(dist_data)
```

The plot below shows the relationship between similarity and distance when different exponents are used. Note that the smallest exponent of 0.1 creates a more local region than the larger exponents.

```{r}
ggplot(dist_data, aes(x = distance, y = similarity, color = exponent)) +
  geom_line() +
  viridis::scale_color_viridis(option = "inferno", discrete = T)
```

