---
title: "LIME Applied to a Logistic Regression"
author: "Katherine Goode"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_depth: 5
    toc_float: true
    theme: cerulean
    highlight: textmate
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  echo = TRUE,
  message = FALSE,
  warning = FALSE,
  eval = TRUE,
  cache = FALSE,
  fig.width = 12
)
```

This journal includes LIME examples with a logistic regression model fit to the sine data.

```{r}
library(caret)
library(cowplot)
library(lime)
library(limeaid)
```

# Data

Prepare the sine data

```{r}
# Prepare training and testing data
x_train = sine_data_train[c("x1", "x2", "x3")]
y_train = factor(sine_data_train$y)
x_test = sine_data_test[1:5, c("x1", "x2", "x3")]
```

# Logistic Regression Models

## Main Effects Only

Logistic regression model: y ~ x1 + x2 + x3

```{r}
# Fit the model
lrm <- caret::train(
  as.factor(y) ~ x1 + x2 + x3,
  preProcess = "scale",
  data = sine_data_train,
  method = "glm",
  family = "binomial"
)

lrm$finalModel
```

## All Interactions 

Logistic regression model: y ~ x1 * x2 * x3

```{r}
# Fit the model
lri <- caret::train(
  as.factor(y) ~ x1 * x2 * x3,
  preProcess = "scale",
  data = sine_data_train,
  method = "glm",
  family = "binomial"
)

lri$finalModel
```

# Apply LIME

Apply LIME with 3 quantile bins, 4 quantile bins, and kernel density to both models

```{r}
# Run apply_lime
resm <- apply_lime(
  train = x_train,
  test = x_test,
  model = lrm,
  label = "1",
  n_features = 2,
  sim_method = c('quantile_bins',
                 'kernel_density'),
  nbins = 3:4,
  return_perms = TRUE
)

# Run apply_lime
resi <- apply_lime(
  train = x_train,
  test = x_test,
  model = lri,
  label = "1",
  n_features = 2,
  sim_method = c('quantile_bins',
                 'kernel_density'),
  nbins = 3:4,
  return_perms = TRUE
)
```

# Exlanation Visualizations

Plot the LIME explanation plot and my EOI plot for each prediction in the test data and each implementation of LIME

```{r}
# Plot the explanation of interest
join_plots <- function(id, exp) {
  p1 <- lime::plot_features(exp$explain[id:(id+1),])
  p2 <- limeaid::plot_explain_scatter(exp$explain[id:(id+1),])
  plot_grid(p1, p2)
}
```

## Main Effects Only 

### 3 quantile bins

```{r}
purrr::map(.x = seq(1, 9, 2), .f = join_plots, exp = resm)
```

### 4 quantile bins

```{r}
purrr::map(.x = seq(11, 19, 2), .f = join_plots, exp = resm)
```

### kernel density

```{r}
purrr::map(.x = seq(21, 29, 2), .f = join_plots, exp = resm)
```

## All Interactions

### 3 quantile bins

```{r}
purrr::map(.x = seq(1, 9, 2), .f = join_plots, exp = resi)
```

### 4 quantile bins

```{r}
purrr::map(.x = seq(11, 19, 2), .f = join_plots, exp = resi)
```

### kernel density

```{r}
purrr::map(.x = seq(21, 29, 2), .f = join_plots, exp = resi)
```